# Getting and clanning data
### Quiz 4, Week 4
============================

Libraries needed and names of the files needed.

```{r}
install.packages("quantmod")
library(quantmod)
file.url<- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
data.q1<- "q1.csv"
file.url2<- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv"
data.q2<- "q2.csv"
file.url4<- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv"
data.q4<- "q4.csv"
```

## Question 1

The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from [here](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv):

```{r, cache=TRUE}
if(file.exists(data.q1)==FALSE){
        download.file(file.url, destfile="q1.csv", method="curl")
}
q1<- read.csv("q1.csv")
head(q1)
```

Apply strsplit() to split all the names of the data frame on the characters "wgtp". **What is the value of the 123 element of the resulting list?**

```{r}
answ1<- strsplit(names(q1), "wgtp")
answ1[123]
```

=============================================

## Question 2

Load the Gross Domestic Product data for the 190 ranked countries in this [data](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FGDP.csv) set:

```{r, cache=TRUE}
if(file.exists(data.q2)==FALSE){
        download.file(file.url2, destfile="q2.csv", method="curl")      
}
q2<- read.csv("q2.csv")
head(q2)
```

Remove the commas from the GDP numbers in millions of dollars and average them. 
**What is the average?**

```{r}
newq2<- q2[5:235,]
newq2<- newq2[1:190,]
names(newq2)<- c("ID_C", "Ranking", "NA", "Country", "GDP", 
                 "NA", "NA", "NA", "NA", "NA")
newq2$GDP<- gsub(",", "", as.character(newq2$GDP))
mean(as.numeric(newq2$GDP))
```
========================================================

## Question 3

In the data set from Question 2 **what is a regular expression that would allow you to count the number of countries whose name begins with "United"? **Assume that the variable with the country names in it is named countryNames. How many countries begin with United?

```{r}
q3<-grep("^United",newq2$Country)
q3
```
================================

## Question 4

Load the Gross Domestic Product data for the 190 ranked countries from before.

Load the educaitional data from this [data](https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FEDSTATS_Country.csv) set: 
Load the educational data from this data set: 

```{r}
if(file.exists(data.q4)==FALSE){
        download.file(file.url4, destfile="q4.csv", method="curl")
}
 q4<- read.csv("q4.csv")
head(q4)
```

Match the data based on the country shortcode. Of the countries for which the end of the fiscal year is available, **how many end in June?**

```{r}
q4$CountryCode<- as.character(q4$CountryCode)
temp<-intersect(q4$CountryCode, newq2$ID_C)
newq4<-data.frame()
for (i in 1:length(temp)){
        temp2<- q4[q4$CountryCode==temp[i],]
        newq4<-rbind(newq4,temp2)
}
answq4<-grep("Fiscal year end: June", newq4$Special.Notes)
length(answq4)
```
==================================================

## Question 5

How many values were collected in 2012? How many values were collected on Mondays in 2012?

```{r}
amzn = getSymbols("AMZN",auto.assign=FALSE)
date<-index(amzn) ## crating a list with the values from Index
year<-grep("*2012", date) ## Create vector with the 2012 values
year<- date[year] ## Create list with only 2012 values
length(year)
days<-weekdays(year) ## Converting into days of the week
sum(days=="Monday") ## Calculate how many mondays.
```
